### 2022-12-14 code/1_exclusions_old/1_exclusionsUS.Rmd
setting up cogsci paper

r_refs("r-references.bib")
r_refs("references/packages.bib")

@@

# Introduction 
By virtue of discrete words and grammatical features, language provides a categorical partition of our continuous experiences. Previous research has observed variation in the cognition of speakers of different languages corresponding to variation within semantic domains in these languages. This language-aligned variation has been observed in the domains of color [@Winawer2007], number [@Frank2008], and gender [@Boroditsky2003], among many others [@Wolff2011]. Differences in cognition between speakers of different languages have often been used to suggest, explain, or exemplify cultural differences between speakers of those languages. Findings in studies of both language and of culture provide convergent evidence suggesting that variation across languages and cultures may shape broad variation in how people think about similarity between, for instance, different colors or objects.

Looking beyond this domain-specific variation, we can also consider broader notions of similarity and relatedness between concepts across domains. The Free Association Norms [@FreeAssoc], for example, measures human judgment in semantic relatedness across domains and contexts. With the availability of large linguistic corpora, we can examine the similarity of concepts that are not circumscribed in individual domains (as color, number, and gender are), allowing us to answer more general questions about variation in cognition between speakers of different languages. By measuring the similarity between arbitrary words (as commonly done with lexical co-occurrence models), it may be possible to model the shape of similarity space in order to capture cross-linguistic and cross-cultural variation and provide a more general answer to the question of how language influences our conception of the world.

Taxonomic and thematic similarity provide a convenient entry point to broader debates about cross-linguistic and cross-cultural variation in notions of similarity. Taxonomic categorization is based on the similarity of attributes, for example, similar perceptual properties, like shared color or shape, among objects. In contrast, thematic categorization is based on causal, spatial, and temporal relationships among objects [@Markman1984]. To illustrate this point, if a participant is given the words “chicken” and “grass”, and asked which of those words are more closely related to the word “cow”, the pairing “chicken” - “cow” indicates a taxonomic match, whereas “grass” - “cow” indicates a thematic match.

@Ji2004 found that preferences for taxonomic and thematic matches differ across culture, even when controlling for testing language and fluency. Specifically, Chinese participants made more thematic matches than taxonomic, while the converse was true for European Americans. This difference held even for Chinese participants tested in English (rather than Mandarin), suggesting that culture has an effect on the categorization strategy independent of the test language. However, the study also found an effect of language within Chinese participants from mainland China (when tested both in China and in the US): their preference for thematic categorization was significantly lower when tested in English compared to Mandarin Chinese. This suggests that test language also influences categorization strategies. Ji et al. characterize these results as demonstrating that culture (independent of language) leads to different styles of reasoning about similarity, whereas language serves as a “tuning” mechanism operating within a culturally-specific style, by activating representations corresponding to the language being used.

## Cross-cultural variation in similarity
The account that culture drives differences in categorization style has been one favored by other non-linguistic studies that use taxonomic vs thematic categorization to probe similarity judgment. A similar study to the aforementioned shows that when shown pictorial triads that represent words (i.e. pictures of an ant, a bee, and honey), Chinese children (9-10 years old) scored higher on thematic categorization compared to American children of the same age range, and vice versa for taxonomic categorization [@Chiu1972]. Additionally, the cross-cultural differences in similarity judgment seem to persist even when the stimuli are more perceptual and less conceptual. @Norenzayan2002 conducted a study where participants are shown a target stimulus beneath two groups of 4 similar objects. The target shares a common feature with all the objects in one group, and only shares another one feature with each member of that group. Meanwhile, the target shares many features with 3 out of 4 members of the other group, but does not share any 1 feature in common with the entire group. The study showed that American participants judged the target to be more similar with the former group more often, and vice versa for Chinese participants. Since differences in categorization style persist with novel stimuli (as opposed to previously seen words or pictures of concepts), this suggests that participants from different cultural backgrounds are utilizing different reasoning strategies when making similarity judgments.

Most commonly, researchers account for the differences in similarity judgment described above by differences in analytic vs holistic processing and in attention between Eastern and Western cultures. According to this framework, East Asians attend more to relationships between objects and context and reason more holistically, whereas Westerners attend more to objects and their properties, and reason analytically [@Nisbett2003]. The analytic-holistic cultural difference in reasoning style has been observed in other non-linguistic tasks. For example, in a rod-and-frame test [@Witkin1954], East Asian (Taiwanese Chinese) participants are more influenced by the position of the surrounding frame than European Americans [@Ji2000], suggesting a higher level of sensitivity to context. East Asians (Koreans) have also been demonstrated to make more situational attributions for behaviors than European Americans, who make more dispositional attributions [@Choi1999], again suggesting higher sensitivity to context. When shown a complex scene with foreground and background information, Japanese participants made more statements about contextual information and relationships between objects than Americans did [@Masuda2001]. Non-linguistic tasks have also shown that participants from Eastern and Western cultures differ in where they focus their visual attention. Another study presented American and Chinese participants with photographs of an object on a complex background [@Chua2005]. They found that Chinese participants made more eye saccades to the background compared to American participants, and that American participants looked at the object more quickly. These cross-cultural differences in attention and analytic vs holistic processing are seen to map onto the cross-cultural differences in similarity judgment: because East Asian participants attend more to the context between objects and reason more holistically, it stands to reason that they would consider words that appear in the same context (i.e. thematically similar) to be more similar than taxonomically-similar words.

Studies that explain these cross-cultural differences in reasoning and similarity judgment with a cultural framework often draw on social and historical contexts as the distal cause for these differences. For example, social structure and social practice can explain why different cultures develop different reasoning habits. To demonstrate this, researchers have looked at regions that are geographically close but with different social structures. @Kitayama2006 showed that Hokkaido Japanese demonstrated greater dispositional bias in attribution of behaviors compared to non-Hokkaido Japanese. The study suggested that this difference is correlated to the fact that Hokkaido was settled quite recently (mid-19th century) by pioneers from southern Japan, and thus its social practices (e.g. daily routines, child-rearing practices, and explicit education) would be influenced by the independent, goal-oriented characteristic of pioneer life. Another study showed that Northern Italians (more independent) categorized objects in a more taxonomic manner than Southern Italians (more interdependent) [@Knight2007]. These studies, however, stopped short of providing a mechanism for the nature of the relationship between differences in social structures and differences in reasoning and similarity judgments.

Another component of the social context explanation seems to be more tenable regarding the causal connection between cross-cultural differences in social practice and those in similarity judgments: that different patterns of attention are passed on by caregivers to children. @Fernald1993 found that American mothers labeled toys and pointed out their attributes more often than do Japanese mothers, while Japanese mothers tended to use the toys to engage their infants in social routines more often than American mothers. @Tardif1997 found that English-speaking caregivers produced more nouns than verbs in child-directed speech, while Mandarin-speaking caregivers emphasized verbs over nouns.  Emphasis on labeling objects and focus on nouns over verbs might lead infants to attend to objects and their taxonomic categorizations, while emphasis on social practices and producing more verbs might direct infants' attention to the relationship between objects. Such differences in early-childhood environments could tune children’s attention to either object or context, and result in cross-cultural differences in similarity judgments (preferring either taxonomic or thematic, respectively) in adulthood.

The above view explains cross-cultural differences in similarity judgment tasks as variation in the conceptualization of similarity itself. Alternatively, these judgments could be shaped by cross-cultural differences in the statistics of the environment, and accordingly the content of everyday experiences. Perhaps when confronted with the triad task, participants from all cultures follow the same process for reasoning about similarity, but rely on language or culture-specific inputs to this process. For example, it may be that Vietnamese participants are more likely to have knowledge of farming practices than Americans, and may associate cows and grass more strongly (as grass is the typical feed for cows in the small-scale family farming contexts common in Vietnam). Americans, on the other hand, may have strong associations between cows and chickens as a result of encountering them together in children’s books or petting zoo contexts. If we observe a difference in categorization between, say Vietnamese and US participants, perhaps they use the same procedure (considering similarity that is influenced by both taxonomic and thematic relations), but the inputs to this procedure differ between cultures, with Vietnamese participants exposed to more instances of “cow” and “grass” occurring together, compared to US participants.

A study by @Miyamoto2006 lends support to this theory. The study randomly sampled scenes from different-sized cities in Japan and the US. Both subjective (judgment by American and East Asian college students) and objective (number of objects as counted by an automated program) showed that Japanese scenes were more complex and contained a larger number of objects than American scenes. This suggests that people living in Japan would more frequently encounter perceptual environments that contain more relationships between objects compared to those who live in the US. Japanese scenes therefore might encourage focus on the overall context, while American scenes might encourage focus on a few salient objects.

Differences in environmental statistics extend beyond the immediate environment to cultural artifacts. @Tsai2007 showed that best-selling storybooks in the US contain more pictures of excited states (e.g. open-mouthed smiles, wide-open eyes) compared to best-selling storybooks in Taiwan, which contain more pictures of calm states. European American preschoolers also perceived excited states as happier and more desirable compared to calm states, and vice versa for Taiwanese Chinese preschoolers. The study also showed that there is a causal link between exposure to environmental statistics and personal preference: regardless of culture, exposure to exciting (vs. calm) storybooks increased children’s preference for the  corresponding activities. Broadly, this suggests that cultural artifacts or information in language (rather than direct, first-hand lived experiences) could be crucial factors in driving cross-cultural variation.

## Language as a proxy of variation in experience
While co-occurrences in experience are difficult to measure, co-occurrence in language can provide a rough proxy--and indeed, language may afford many of the “experiences” that people have with infrequently encountered items, like cows or helicopters. In the planned analyses, we will use co-occurrence in language as a proxy of cultural experience. This is a generalizing assumption, and while no linguistic corpus can be expected to fully capture a culture, and different linguistic corpora may capture different variables of a culture, using this proxy as a model of culture provides a conservative test of the hypothesis that differences in the input, rather than the process, of reasoning produces cultural variation in this task. This means that if such a model does work -- if linguistic co-occurrence can predict cultural variation in this similarity task without building differences in the decision process into the model -- it is relatively strong evidence that language alone (perhaps as a more easily observed source of information on broader cultural and ecological variation) may account for these cross-cultural differences in reasoning.[^footnote]

Indeed, previous work suggests that using lexical co-occurrence as a proxy can be useful in thinking about similarity judgments. Natural language processing tasks have found that lexical co-occurrence is a good predictor for performance in similarity judgment tasks. @Griffiths2007 showed that a model that takes word-document co-occurrence as input can be used to predict word association and the effects of semantic association on a variety of linguistic tasks. Additionally, lexical semantic models that are constructed using lexical co-occurrence (as opposed to annotated relations) have been shown to perform well on predicting human judgments about similarity between word pairs that are either thematically or taxonomically related [@Rohde2006].

Our current study will be using 2 measures of lexical co-occurrence: 1) raw lexical co-occurrences and 2) cosined distances of fastText word vectors. fastText is a system that uses lexical co-occurrence information to generate a vector representing each word in its lexicon [@Mikolov2018]. fastText uses a continuous bag-of-words model, which takes into account a symmetric window of words surrounding the word in question and maximizes the log-likelihood of the probability of the word given its context. Calculating the cosine-similarity between word vectors from fastText gives us a score of how similar the contexts of these words are. Therefore, words that occur in the same context (high lexical co-occurrence) would have a larger cosine-similarity score.

It must be noted that fastText supplements the lexical co-occurrence information by simultaneously training weight vectors for each position relative to the word in question, as well as adding sub-word information (bag-of-character n-gram vectors for each character forming the words). The similarity judgment prediction given by fastText may be more higher than a model using purely lexical co-occurrence information. However, models that use only lexical co-occurrence information (with or without a simple position weight system that gives more weight to words that are closer to the word in question) also perform well in similarity prediction tasks. For this study, due to the lack of large cleaned corpora for Mandarin and Vietnamese, we are leveraging fastText pre-trained models on the languages we are interested in @Grave2018.

Systems like fastText and word2vec have been demonstrated to be good predictors for similarity judgments. @Liu2019 showed that fastText performed well as the data for a system matching responses in a knowledge base by how similar they are to a customer’s query. A study by @Jatnika2019 tested a word2vec model trained on the English wikipedia corpus against 353 word pairs labeled with similarity values based on human judgment. The word2vec model correlates moderately with human judgment. fastText has also been shown to be sensitive towards cultural effects on word meanings: @Thompson2020 applied fastText to language-specific Wikipedia corpus (among others) to generate a “semantic neighborhood” of 1010 meanings in different languages. The study showed that languages that are spoken by more similar cultures, more geographically proximate and/or historically related had more semantic neighborhoods that aligned. This is a proof of concept for our use of fastText in different languages with varying levels of relatedness.

## The present study
@Ji2004 showed the difference in similarity judgment between Chinese and European Americans (the former prefers thematic categorization while the latter prefers taxonomic). However, their evidence is not sufficient to conclude that it is culture that drives this difference, as their cross-cultural data is correlational. Furthermore, they have yet to rule out all mechanisms by which language could be relevant to the task. For example, their language manipulation (bilingual speakers from Mainland China making more taxonomic categorizations when tested in language) only points to an offline effect of language because it does not prevent participants from accessing linguistic representations. A stronger claim could have been made if verbal interference had been conducted, and if cross-cultural differences had remained. This would have been stronger in eliminating online effect of language as a potential cause for cross-cultural differences in similarity judgment.

Additionally, outside of language, cross-cultural differences could come from a couple different mechanisms: @Ji2004 suggested that the notion of similarity is what varies across cultural groups, but it’s quite possible that similarity is regarded in the same way and it’s actually the input to this similarity judgment that varies across culture contexts. It might be possible to gain traction on possible mechanisms by examining whether variation in similarity judgments co-varies with environmental statistics that differ across the cultural and linguistic contexts.

In this study, we ask whether variation in lexical co-occurrence can account for the cross-context (culture and language) differences observed in how people evaluate similarity. Firstly, we attempt to replicate @Ji2004 by running a study comparing English speakers in the US and Mandarin speakers in mainland China. Secondly, we evaluate whether the differences observed between English and Mandarin speakers’ similarity judgments also extend to comparisons between English and Vietnamese, by running a similar study between English speakers in the US and Vietnamese speakers in Vietnam. Vietnam is a Southeast Asian country that borders China and is historically greatly influenced by Chinese culture [@Hui2002]. Therefore, it serves as a suitable cultural context to investigate whether the claim made by @Ji2004 and previous studies, that Eastern and Western cultures have different notions of similarity, extends beyond mainland China.

We then ask to what extent similarity judgments show cross-context differences that align with co-occurrence patterns in participants’ languages. We would only see a main effect of cultural context if the notion of differing notions of similarity is driving these cross-cultural differences. Meanwhile, if we do see variation in responding trial-to-trial that tracks with lexical co-occurrence statistics (as a rough proxy combining both linguistic and cultural context), this would give traction to the view that cross-cultural notions of similarity might be similar, and it is environmental statistics that guide the differences in similarity judgments. To preview our results, we find that, while we do not see overall group differences (between the East Asian contexts and the US) in similarity, we do find language-specific co-occurrence can indeed explain cross-context differences in similarity judgments. In contrast to previous accounts, by which culture induces differing conceptions of similarity (varying between taxonomic and thematic), these findings provide an alternative, and more specific, account by which language may explain these cross-cultural differences without invoking variation in notions of similarity.

# Methods 
```{r get exclusion results from 1_exclusions.Rmd}
source("../1_exclusions/1_exclusionsCN.Rmd")
source("../1_exclusions/1_exclusionsUS.Rmd")
source("../1_exclusions/1_exclusionsVN.Rmd")
```

```{r demog_E, TODO: move outside, include = F}
demog <- c("demog-dropdown", "demog-multi-choice", "demog-free-response", "demog-multi-select")
df.demog_E <- df.resp_E %>%
  filter(trial_type %in% demog) 
#write.csv(data_demog, "data_anon_demog_0201.csv")

AGE = 97 #index for age answer
GENDER = 115 #index for gender answer
split_responses <- function(responses) {
  responses <- as.character(responses)
  response_list <- unlist(strsplit(responses, ',|:'))
  response_list <- response_list[seq(1, length(response_list))]
  response_list <- gsub("}", "", response_list)
  response_list <- gsub("\"", "", response_list)
  return(response_list)
}

get_age <- Vectorize(function(responses) {
  responses <- split_responses(responses)
  if(responses[AGE-1] != "A0") {warnings("cannot find age answer")}
  else{return (responses[AGE])}
})

get_gender <- Vectorize(function(responses) {
  responses <- split_responses(responses)
  if(responses[GENDER-1] != "A1") {warnings("cannot find gender answer")}
  else{return (responses[GENDER])}
})

df.age_gender_E <- df.demog_E %>%
  filter(variable_type == "demog_age_gender_ethnic") %>%
  select(subject, responses) %>%
  mutate(age = as.numeric(get_age(responses))) %>%
  mutate(age = as.numeric(age)) %>%
  mutate(gender = get_gender(responses))

df.age_gender_E_summ <- df.age_gender_E %>%
  summarize(mean_age = mean(age), 
            median_age = median(age), 
            sd_age = sd(age))
```

```{r demog_V, TODO: move outside, include = F}
demog <- c("demog-dropdown", "demog-multi-choice", "demog-free-response", "demog-multi-select")
df.demog_V <- df.resp_V %>%
  filter(trial_type %in% demog)

AGE = 97 #index for age answer
GENDER = 115 #index for gender answer
split_responses <- function(responses) {
  responses <- as.character(responses)
  response_list <- unlist(strsplit(responses, ',|:'))
  response_list <- response_list[seq(1, length(response_list))]
  response_list <- gsub("}", "", response_list)
  response_list <- gsub("\"", "", response_list)
  return(response_list)
}

get_age <- Vectorize(function(responses) {
  responses <- split_responses(responses)
  if(responses[AGE-1] != "A0") {warnings("cannot find age answer")}
  else{return (responses[AGE])}
})

get_gender <- Vectorize(function(responses) {
  responses <- split_responses(responses)
  if(responses[GENDER-1] != "A1") {warnings("cannot find gender answer")}
  else{return (responses[GENDER])}
})

df.age_gender_V <- df.demog_V %>%
  filter(variable_type == "demog_age_gender_ethnic") %>%
  select(subject, responses) %>%
  mutate(age = as.numeric(get_age(responses))) %>%
  mutate(age = as.numeric(age)) %>%
  mutate(gender = get_gender(responses))

df.age_gender_V_summ <- df.age_gender_V %>%
  summarize(mean_age = mean(age), 
            median_age = median(age), 
            sd_age = sd(age))
```

```{r demog_M, TODO: MOVE outside, include = F}
demog <- c("demog-dropdown", "demog-multi-choice", "demog-free-response", "demog-multi-select")
df.demog_M <- df.resp_M %>%
  filter(trial_type %in% demog) 
#write.csv(data_demog, "data_anon_demog_0201.csv")

AGE = 97 #index for age answer
GENDER = 115 #index for gender answer
split_responses <- function(responses) {
  responses <- as.character(responses)
  response_list <- unlist(strsplit(responses, ',|:'))
  response_list <- response_list[seq(1, length(response_list))]
  response_list <- gsub("}", "", response_list)
  response_list <- gsub("\"", "", response_list)
  return(response_list)
}

get_age <- Vectorize(function(responses) {
  responses <- split_responses(responses)
  if(responses[AGE-1] != "A0") {warnings("cannot find age answer")}
  else{return (responses[AGE])}
})

get_gender <- Vectorize(function(responses) {
  responses <- split_responses(responses)
  if(responses[GENDER-1] != "A1") {warnings("cannot find gender answer")}
  else{return (responses[GENDER])}
})

df.age_gender_M <- df.demog_M %>%
  filter(variable_type == "demog_age_gender_ethnic") %>%
  select(subject, responses) %>%
  mutate(age = as.numeric(get_age(responses))) %>%
  mutate(age = as.numeric(age)) %>%
  mutate(gender = get_gender(responses))

df.age_gender_M_summ <- df.age_gender_M %>%
  summarize(mean_age = mean(age), 
            median_age = median(age), 
            sd_age = sd(age))
```
## Participants
We recruited `r length(unique(df.ppts_finished_E$subject))` participants from the US, `r length(unique(df.ppts_finished_V$subject))` participants from Vietnam, and `r length(unique(df.ppts_finished_M$subject))` participants from mainland China. US participants were recruited through snowball sampling seeded with Stanford email lists, Vietnam participants were recruited through snowball sampling seeded with Vietnam-based student groups on Facebook, and mainland China participants were recruited through snowball smapling seeded with group chats on WeChat. US participants were compensated with $5 gift certificates (USD), VN participants received 50,000₫ (VND) in phone credit, and mainland China participants received 25CNY through WeChat credit transfer.

We excluded `r length(unique(df.ppts_finished_E$subject)) - length(unique(df.ppts_att_E$subject))` US participants, `r length(unique(df.ppts_finished_V$subject)) - length(unique(df.ppts_att_V$subject))` Vietnam participants, and `r length(unique(df.ppts_finished_M$subject)) - length(unique(df.ppts_att_M$subject))` China participants who did not answer all attention checks correctly. We followed 4 exclusion criteria that aim to retain only participants who are influenced by one culture: (1) non-native speakers of English and Vietnamese, respectively, (2) fluent in at least one of the other two study languages (Vietnamese for US participants, English for Vietnamese participants and Chinese participants), (3) have lived outside of the test country (US, Vietnam, or China) for more than two years, and (4) have significant international experience (more than 6 international experiences of 2 days or longer.) We did not use a particular criterion for a language if it would exclude 25% or more of any one sample. In this round of exclusion, we excluded `r length(unique(df.ppts_att_E$subject)) - length(unique(df.resp_E$subject))` US, `r length(unique(df.ppts_att_V$subject)) - length(unique(df.resp_V$subject))` Vietnam participants, and `r length(unique(df.ppts_att_M$subject)) - length(unique(df.resp_M$subject))` China participants. After these exclusions, the US sample included `r length(unique(df.resp_E$subject))` participants (`r df.age_gender_E %>% filter(gender == "Male") %>% count()`M, `r df.age_gender_E %>% filter(gender == "Female") %>% count()`F, `r df.age_gender_E %>% filter(gender == "Non-binary") %>% count()` non-binary, `r df.age_gender_E %>% filter(gender == "Decline to answer") %>% count()` other), with mean age = `r round(df.age_gender_E_summ$mean_age, 2)` (SD = `r round(df.age_gender_E_summ$sd_age, 2)`) and median age = `r df.age_gender_E_summ$median_age`. The Vietnam sample included `r length(unique(df.resp_V$subject))` participants (`r df.age_gender_V %>% filter(gender == "Nam") %>% count()`M, `r df.age_gender_V %>% filter(gender == "Nữ") %>% count()`F), with mean age = `r round(df.age_gender_V_summ$mean_age, 2)` (SD = `r round(df.age_gender_V_summ$sd_age, 2)`) and median age = `r df.age_gender_V_summ$median_age`. The China sample included `r length(unique(df.resp_M$subject))` participants (`r df.age_gender_M %>% filter(gender == "男性") %>% count()`M, `r df.age_gender_M %>% filter(gender == "女性") %>% count()`F), with mean age = `r round(df.age_gender_M_summ$mean_age, 2)` (SD = `r round(df.age_gender_M_summ$sd_age, 2)`) and median age = `r df.age_gender_M_summ$median_age`.

Notably, we lost a majority of our Vietnam participants (more than 60%) in the attention check exclusion. One reason we suspect why this might have happened is because our Vietnam participants are less familiar with research surveys and attention check questions, and thus might have thought too much about the attention check questions. We carried out an exploratory analysis where we used a less stringent attention check exclusion (covered in the final part of the Results & Analysis section). To preview our findings for this analysis, we did not find any difference in significant results when compared with the main analysis.

## Stimuli
We adapted stimuli from previous studies to create a set of test triads consisting of a cue, with one thematic and one taxonomic match option. For example, “cow,” “grass,” and “chicken,” where “cow” is the cue, “grass” is the thematic match, and “chicken” the taxonomic match. We included 105 such triads, a superset including triads pulled from supplemental information and in-text examples across the literature, and others that we adapted or created (see SI for the full list of stimuli and sources). We selected triads on the basis of cultural familiarity in the US, Vietnam, and China. The triads were originally in English; they were translated to Vietnamese and Mandarin by a fluent bilingual speaker in each language. The translations were then checked for accuracy after backtranslation to English by another fluent bilingual in each language who was naive to the original English versions.

Each participant completed all 105 triads in sets of 21 trials at a time (10 test triads, 10 filler triads, and 1 attention check per page), by selecting the match most related to the cue (“Which thing is most closely related to the bolded item?” or "Thu nào liên quan nhat voi thu đuoc in đam?"). The test triads were presented with 105 filler triads mixed in, to obscure the taxonomic-thematic two-answer forced choice structure of the test stimuli and reduce the likelihood that participants would become aware of the design. The filler triads were groups of three semantically related words, but where the match options were not distinguished by thematic vs. taxonomic similarity, e.g., cue “bird” with match options “lizard” and “toad.” Additionally, we included 10 attention check trials, which were formatted like the test and filler triads but included an instruction instead of a cue item, e.g., “Choose wife” with match options “wife” and “husband.” In total, each participant completed 210 similarity judgments and 10 attention check questions, with triads presented in randomized orders that varied between subjects.

## Corpus model 
Our general approach is to build a model of similarity that is based on collocation counts in each language.

### Raw lexical co-occurrences
To give an intuition for our model, consider the cow-grass-chicken triad: we counted how many times “cow” and “grass” co-occur within a window of text in each corpus, and compare this to how many times “cow” and “chicken” co-occur. Our similarity prediction is then proportional to the relative frequency of these pairs. For example, if the thematic cow-grass match accounts for 30% of collocations for this triad (with cow-chicken making up the other 70%), then our model predicts, correspondingly, that 30% of responses to the triad will be grass, and the other 70% chicken. We then use a mixed-effects regression to evaluate how well each corpus collocate model predicts participants’ similarity judgments, across triads and languages.

#### Collocate retrieval and coding
For our English co-occurrence metric, we used the online interface of the Corpus of Contemporary American English [@COCA] to retrieve collocation counts. We recorded the raw count of times that any cue-match pair (e.g., cow-grass or cow-chicken) co-occur in a window of 19 words, which is the maximum window size in the online interface, and closest to the sentence co-occurrence metric in our VI corpus collocation counts.

To determine Vietnamese co-occurrence, we used the raw frequency of sentence co-occurrences from a subset of the Vietnamese corpus in the Leipzig Corpora Collection [@VICorpus]; this corpus includes 70 million Vietnamese sentences, but our corpus data comes from a 1 million sentence subset for which co-occurrence counts are available for download. Vietnamese makes very frequent use of compositional morphology but the written language  uses spacing to delineate syllable boundaries rather than word boundaries. Accordingly, collocate searches returned instances of both target terms and many morphologically related, but distinct, words. We included in our counts any instances of the target term or close semantic neighbors containing the same morpheme(s) as long as they entailed a likely literal reference to the target term. For example, our search for collocates of the term “gà” (chicken) also returned “gà mái,” a distinct word for female chickens. Despite being a different word, “gà mái” both includes the morpheme for “chicken” and entails reference to a chicken. Accordingly, instances of both “gà” and “gà mái” were included in our collocate count for “chicken.” Some compounds do not meet this criteria, and were excluded from collocate counts. For example “trái cây” (tree.fruits) includes the same syllable as the word “cây” (tree), but refers to fruit that comes from trees, not the trees themselves, and was therefore excluded from our collocate counts for “tree.”

We were unable to obtain a Mandarin corpus with similar number of sentences as the aforementioned English and Vietnamese corpora, hence we did not include any analysis based on raw lexical co-occurrences with Mandarin.

#### Collocate similarity model
From the raw co-occurrence counts of each triad, we calculated the thematic co-occurrence ratio as the number of thematic co-occurrences over the sum of thematic and taxonomic co-occurrences. We did this for both the English and Vietnamese corpus co-occurrence counts. In this way, we obtained predictions for 73 of 105 triads from the Vietnamese corpus and 104 of 105 triads from English. We therefore limited all analyses reported here to the subset of triads for which our corpus-based model can make meaningful predictions, meaning triads that have at least one non-zero collocate count (either thematic or taxonomic match). We replaced any remaining zero collocate counts with $\epsilon$ to account for sparsity in the corpus data. We then tested whether these simple relative frequency models predict US and VN responding.

### Cosine distance of word vectors
To give an intuition for our model, consider again the cow-grass-chicken triad: we retrieved word vectors for "cow" and "grass", and calculate the cosine distance between these vectors. Similarly, we retrieved vectors for "cow" and "chicken" and calculate the cosine distance between them. Our similarity prediction is then inversely proportional to the cosine distance of these pairs. This is because a larger cosine distance means the word vectors are further apart, and thus the words are less similar. For example, if the cosine distance of thematic cow-grass is 0.7 and the cosine distance of taxonomic cow-chicken is 0.3, then our model predicts, correspondingly, that 30% of responses to the triad will be grass, and the other 70% chicken. We then use a mixed-effects regression to evaluate how well each corpus model predicts participants’ similarity judgments, across triads and languages.

#### Collocate retrieval
We use the fastText pre-trained models of English, Mandarin, and Vietnamese in @Grave2018. These models are trained on Common Crawl and Wikipedia using We distribute pre-trained word vectors for 157 languages, trained on Common Crawl and Wikipedia using fastText. These models were trained using a Continuous Bag of Words (CBOW) with position-weights and a window of size 5. The models use character n-grams of length 5 and 10  negative examples.From the aforementioned models, we retrieve the word vectors (dimension 300) for each word we are interested in.

#### Collocate similarity model
From the word vectors, we calculated the cosine distance between each cue-thematic match (thematic cosine distance) and cue-taxonomic match (taxonomic cosine distance), using the spatial.distance.cosine function from the SciPy package [@2020SciPy]. We then calculated the thematic cosine distance proportion as thematic cosine distance over the sum of taxonomic cosine distance and thematic cosine distance. We did this for all three corpora. We were able to obtain predictions for all triads in all languages.

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results
```{r combine corpus and behavior data, include = F}
#source("../2_combine/2_combine.Rmd")
df <- read.csv("../../data/data_USCNVN_ENZHVI.csv")
```

## Do we extend previous work reporting a preference for taxonomic matching in the US and thematic matching in Asia?

```{r warning=FALSE, include=FALSE, paged.print=TRUE}
df.country <- df %>%
  group_by(subject, country) %>%
  summarize(theme_resp_percent = mean(responses_theme, na.rm = T))

df.country_sum <- df.country %>%
  group_by(country) %>%
  summarize(mean_theme_resp_percent = mean(theme_resp_percent), 
            sd_theme_resp_percent = sd(theme_resp_percent))

fit.country = glmer(responses_theme ~ country + (1 | subject) + (country | triad), 
                    data = df, 
                    family = "binomial")
summary(fit.country) 
fit.country.anova = Anova(fit.country, type=3)

fit.country_EN_VN = glmer(responses_theme ~ country + (1 | subject) + (country | triad), 
                    data = df %>% filter(country != "China"),
                    family = "binomial")
summary(fit.country_EN_VN) 
```
The group means of proportion of thematic response in mainland China is the highest (M = `r round((df.country_sum %>% filter(country == "China"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "China"))$sd_theme_resp_percent, 2)`), followed by the groups means in Vietnam (M = `r round((df.country_sum %>% filter(country == "Vietnam"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "Vietnam"))$sd_theme_resp_percent, 2)`), which is slightly higher than that of the US (M = `r round((df.country_sum %>% filter(country == "US"))$mean_theme_resp_percent, 2)`, SD = `r round((df.country_sum %>% filter(country == "US"))$sd_theme_resp_percent, 2)`).

```{r echo=FALSE, warning=FALSE, fig.cap="Proportion of thematic responses by country. Only US-China responding comparison shows a siginficant difference. We could not extend this to US-Vietnam responding comparison."}

#show violin plot
ggplot(df.country,
       mapping = aes(x = factor(country, levels=c("China", "US", "Vietnam")), 
                     y = theme_resp_percent, 
                     color = factor(country, levels=c("China", "US", "Vietnam")))) +
  geom_violin() +
  geom_jitter(height = 0, 
              alpha = 0.3) +  
  stat_summary(fun.data = "mean_cl_boot", 
               geom = "pointrange") + 
  labs(y = "Proportion Thematic Chosen", 
       x = "Country", 
       color = "Country") + 
  scale_color_manual(values=c("#D63230", "#1C77C3", "#F39237"))
```

```{r correlation of country response, echo=FALSE, warning=FALSE}
#US-China, US-Vietnam, China-Vietnam
df.triad_country <- df %>%
  group_by(triad, country) %>%
  summarize(emp_theme_prop = mean(responses_theme))

df.triad_country <- pivot_wider(data = df.triad_country, 
                                names_from = country, 
                                values_from = emp_theme_prop)
plotUSCN = ggplot(data = df.triad_country,
                  mapping = aes(x = US, 
                                y = China, 
                                label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)", 
        y = "Proportion Thematic Chosen (China)") + 
   xlim(0, 1) + 
   ylim(0, 1) + 
  theme(axis.title = element_text(size=rel(0.75)))

plotUSVN = ggplot(data = df.triad_country,
                  mapping = aes(x = US, 
                                y = Vietnam, 
                                label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (US)", 
        y = "Proportion Thematic Chosen (Vietnam)") + 
   xlim(0, 1) + 
   ylim(0, 1)  + 
  theme(axis.title = element_text(size=rel(0.75)))

plotCNVN = ggplot(data = df.triad_country,
                  mapping = aes(x = Vietnam,
                                y = China, 
                                label = triad)) +
   geom_text() + 
   geom_smooth(method = "lm") +
   labs(x = "Proportion Thematic Chosen (Vietnam)", 
        y = "Proportion Thematic Chosen (China)") + 
   xlim(0, 1) + 
   ylim(0, 1)  + 
  theme(axis.title = element_text(size=rel(0.75)))

corr.USCN = apa_print(cor.test(df.triad_country$US, 
                  df.triad_country$China, na.rm=T))

corr.USVN = apa_print(cor.test(df.triad_country$US, 
                  df.triad_country$Vietnam, na.rm=T))

corr.CNVN = apa_print(cor.test(df.triad_country$China, 
                  df.triad_country$Vietnam, na.rm=T))
```

```{r warning = F, fig.width = 8, fig.cap="Correlation between proportion thematic chosen by US-China participants, US-Vietnam participants, and China-Vietnam participants. All three comparisons show strong correlation, suggesting a strong signal of universality across cultural contexts regarding similarity judgments"}
plotUSCN + plotUSVN + plotCNVN +
  plot_layout(ncol = 3)
```

With a Pearson correlation test, we observed high correlations between thematic responding across all three countries. The proportion of US thematic responding is strongly correlated with thematic responding in China, `r corr.USCN$full_result`. The proportion of US thematic responding is strongly correlated with thematic responding in Vietnam, `r corr.USVN$full_result`. Likewise, the proportion of China thematic response is strongly correlated with thematic responding in Vietnam, `r corr.CNVN$full_result`. This points to a strong signal of universality across cultural contexts regarding similarity judgments.

```{r USCN effect size comparison, include=FALSE}
USCN_cohen <- cohen.d(df.triad_country$China, df.triad_country$US)
```

>>>>

@@

```{r loading data E, warning=FALSE, include=F}
df.resp_E <- read.csv("../../data/responding/data_US.csv") %>%
  select(-X)
